---
title: "Analise de Dados e Simulação - Lista 1"
author: "André Vinícius, Kaio"
date: "March 27, 2019"
output: html_document
---

In this first report, we will try our first steps on data analysis with information from https://www.citibikenyc.com/system-data

# libraries used
```{r libraries, render=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggmap)
```


# downloading the data
The original file is in https://s3.amazonaws.com/tripdata/index.html

```{r downloading data}
bikeurl <- "https://s3.amazonaws.com/tripdata/201902-citibike-tripdata.csv.zip"
# here I had first to install curl on Ubuntu
unzippedfile <- "201902-citibike-tripdata.csv"
if (!file.exists(unzippedfile))
    {
    destfile <- "bikesfev.csv.zip"
        if (!file.exists(destfile))
        {
        download.file(bikeurl, destfile = destfile, method = "curl")
        downloadtime <- date()
        }
    unzip(destfile)
    }
dadosbike <- read.csv(unzippedfile)
```
# looking at the data
```{r checking data}
head(dadosbike)
str(dadosbike)
summary(dadosbike)
```
## then I had to tidy some of the data
# Tidy 1: numeric to factor
```{r tidying}
dadosbike$start.station.id <- as.factor(dadosbike$start.station.id)
dadosbike$end.station.id <- as.factor(dadosbike$end.station.id)
dadosbike$bikeid <- as.factor(dadosbike$bikeid)
dadosbike$gender <- as.factor(dadosbike$gender)
# there is no need to do the following steps, but I'll keep them
# as.factor(dadosbike$birth.year)
# Tidy 2: 
# dadosbike <- separate(dadosbike, starttime, c("startdate", "starttime"), " ")
# dadosbike <- separate(dadosbike, endtime, c("enddate", "endtime"), " ")
# Tidy 3: string to date
# dadosbike$startdate <- as.Date(dadosbike$startdate)
# dadosbike$enddate <- as.Date(dadosbike$enddate)
# Tidy 4: string to hrmour(lubridate)
# dadosbike$starthour <- hms(dadosbike$starthour)
# dadosbike$endhour <- hms(dadosbike$endhour)
```
#cleaning the data
How to get machado's 99.5 percentile:
```{r cleaning data}
plot(sort(dadosbike$tripduration), log = "y")
plot(quantile(dadosbike$tripduration, probs = c(0:100)/100))
quantile(dadosbike$tripduration, probs = c(0, 0.005, 0.25, 0.5, 0.75, 0.995, 1))
```
#A lot of ways to explore the data

 tripduration by - station, time, day, usertype\
 start/end time - create faixas\
 weekday, monthday, month, season\
 station by - tripduration(descriptive), birthyear(histogram), usertype, count\
 bikeid?\
 usertype by - station, birthyear, time, day(weekday, monthday, etc), tripduration\
 birthyear by - station, time, day, usertype, tripduration\
 create distance variable?\

# histograms
 wday by subscriber\
 wday by consumer\
 hour by subscriber\
 hour by consumer\
 day/hour peaks\
 all these by station\

#boxplots
 tripduration by subscriber\
 tripduration by consumer\
 birthyear\


# plots
How to plot over the map:\
install ggmap library, read ?register_google, and read https://www.r-bloggers.com/how-to-plot-basic-maps-with-ggmap/


```{r plotting on map}
x <- dadosbike$start.station.longitude
y <- dadosbike$start.station.latitude
mapinfo <- data.frame(x, y)
mapinfo <- unique(mapinfo)
# https://rdrr.io/cran/ggmap/man/make_bbox.html
bbox <- make_bbox(mapinfo$x, mapinfo$y)

get_map(bbox) %>%
  ggmap() +
  geom_point(data = mapinfo,
             aes(x = mapinfo$x,
                 y = mapinfo$y))

start.station.count <- dadosbike %>%
  select(start.station.longitude, start.station.latitude) %>%
  count(start.station.longitude, start.station.latitude)

end.station.count <- dadosbike %>%
  select(end.station.longitude, end.station.latitude) %>%
  count(end.station.longitude, end.station.latitude)

# https://www.littlemissdata.com/blog/maps
get_map(bbox) %>%
  ggmap() +
  geom_point(data = start.station.count,
             aes(x = start.station.longitude,
                 y = start.station.latitude),
             size = (start.station.count$n/2000))

get_map(bbox) %>%
  ggmap() +
  geom_point(data = end.station.count,
             aes(x = end.station.longitude,
                 y = end.station.latitude),
             size = (end.station.count$n/2000))
```

## How to use these:
Try ?geom_path, ?geom_step, ?geom_line


# compare histogram vs. boxplot!!!

##A lot of things I have tried but have to fix the code before plotting
head(dadosbike$starttime)\
dadosbike$starttime <- hms(dadosbike$starttime)\
dadosbike$endtime <- hms(dadosbike$endtime)\
head(dadosbike)\
summary(dadosbike)\
tail(arrange(dadosbike$bikeid))\
tail(arrange(dadosbike$bikeid))\
dadosbike %>% select(bikeid) %>% count(bikeid) %>% arrange(desc(n))\
dadosbike %>% select(bikeid) %>% count(bikeid) %>% arrange(n)\
dadosbike %>% select(bikeid) %>% count(bikeid) %>% summarise(mean(n))\
dadosbike %>% select(bikeid) %>% count(bikeid) %>% summarise(median(n))\
dadosbike %>% select(bikeid) %>% count(bikeid) %>% summarise(boxplot(n))\
boxplot(dadosbike %>% select(bikeid) %>% count(bikeid) %>% select(n))\
boxplot(hour(dadosbike$starttime))\
boxplot(hour(dadosbike$starttime)) ~ dadosbike$start.station.id\
boxplot(hour(dadosbike$starttime)) ~ dadosbike$start.station.id\
boxplot(hour(dadosbike$starttime)) ~ dadosbike$start.station.id\
boxplot(hour(dadosbike$starttime)) ~ dadosbike$start.station.id\
hist(hour(dadosbike$starttime))\
hist(wday(dadosbike$startday))\
dadosbike %>% filter(usertype == "Customer") %>% select(startday)
hist(wday(dadosbike$startday))\
hist(wday(dadosbike$startday[dadosbike$usertype == "Customer"]))\
hist(wday(dadosbike$startday[dadosbike$usertype == "Subscriber"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Customer"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Subscriber"]))\
unique(dadosbike$start.station.name)\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Customer"
                              & dadosbike$start.station.name == "Grove St PATH"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Subscriber"
                              & dadosbike$start.station.name == "Grove St PATH"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Customer"
                              & dadosbike$start.station.name == "Morris Canal"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Subscriber"
                              & dadosbike$start.station.name == "Morris Canal"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Customer"
                              & dadosbike$start.station.name == "Sip Ave"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Subscriber"
                              & dadosbike$start.station.name == "Sip Ave"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Customer"
                              & dadosbike$start.station.name == "Hamilton Park"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Subscriber"
                              & dadosbike$start.station.name == "Hamilton Park"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Customer"
                              & dadosbike$start.station.name == "Journal Square"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Subscriber"
                              & dadosbike$start.station.name == "Journal Square"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Customer"
                              & dadosbike$start.station.name == "JC Medical Centel"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Subscriber"
                              & dadosbike$start.station.name == "JC Medical Center"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Customer"
                              & dadosbike$start.station.name == "Lincoln Park"]))\
hist(hour(dadosbike$starttime[dadosbike$usertype == "Subscriber"
                              & dadosbike$start.station.name == "Lincoln Park"]))\
hist(as.numeric(dadosbike$birth.year))\
hist(as.numeric(dadosbike$birth.year[dadosbike$usertype == "Subscriber"]))\
hist(as.numeric(dadosbike$birth.year[dadosbike$usertype == "Customer"]))\
boxplot(hour(dadosbike$starttime))\
boxplot(as.numeric(dadosbike$birth.year[dadosbike$usertype == "Subscriber"]))\
boxplot(as.numeric(dadosbike$birth.year[dadosbike$usertype == "Customer"]))\
boxplot(hour(dadosbike$starttime)~dadosbike$birth.year)\
boxplot(hour(dadosbike$starttime)~dadosbike$usertype)\
boxplot(hour(dadosbike$starttime[dadosbike$usertype == "Subscriber"]~
               dadosbike$start.station.name))\